### 2013-06-28 Tex/chapters/introduction.tex
Viele kleiner verbesserungen, umformulierengen präzisierungen
weitere Korrekturen und Umformulierungen

Unfortunatly the memory of any computer is small in comparison with the
thermodynamic limit. This leads to \emph{finite size effects}.
In section \ref{ssec:finitesize} will be discussed how to manage them.\\

>>>>

Unfortunately the memory of any computer is small in comparison with the
thermodynamic limit. This leads to deviations from the real behavior of
the system in the thermodynamic limit. These deviations are called
\emph{finite size effects} and in section \ref{ssec:finitesize} will be
discussed how to manage them.\\

### 2013-06-28 Tex/chapters/monte_carlo.tex
weitere Korrekturen und Umformulierungen

\label{sec:montecarlo}
%~ Monte Carlo simulations are (probably) named after the casino in Monaco
%~ \cite{NewmanBarkema1999} because both rely on randomness.\\
The idea behind Monte Carlo simulations is to take random samples of
the observable, which should be measured, and estimate the observable from
this samples. In example the Monte Carlo integration chooses random
points within the bounding box of the fuction and estimates the integral
as the fraction of points below the function and points above muliplied
by the area of the bounding box.
Of course one wants to sample more points, where the area below the
function is big, because points in areas, where the function does
contribute little to the integral, contribute little to the integral.
This can be achieved by \emph{Importance Sampling}. One
generates random numbers according to some probabilty distribution and
weights the result accordingly.\\
%~ To improve this method: \emph{Importance Sampling}
%~ means that one takes more samples where the contribution of the
%~ function to the integral is bigger by using random numbers
%~ distributed........
The same principle is used to sample properties of statistical systems
by generating random states.
But it is difficult to create random state of a physical system i.e. the
Ising system in equilibrium, hence one uses \emph{Markov Chains} to
generate new states \(\nu\) from old ones \(\mu\).
It is important that the transition probabilities \(A(\mu \to \nu)\)
obey \emph{Detailed Balance} and \emph{Ergodicity}.
\emph{Detailed Balance} means  that the probability to leave a state is
the same as the probability to enter the state in equilibrium
\(p_\mu A(\mu \to \nu) = p_\nu A(\nu \to \mu)\) with \(p_\mu\) the
probability to be in state \(\mu\).
And \emph{Ergodicity} requires that every possible state is reachable
from every other state in finite time. \cite{NewmanBarkema1999} \cite{Katzgraber2011}
Otherwise the samples might not be representative for the whole system.\\

\subsection{Monte Carlo Algorithms}
    \paragraph{Metropolis}
        A Metropolis Monte Carlo simulation of an Ising model will
        choose a random spin, calculate the energy change
        \(\Delta \hat H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of that spin and excute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta \hat H = \hat H(\nu) - \hat H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                                 & \Delta \hat H \le 0 \\
                \exp{\brac{-\beta \Delta \hat H}} & \Delta \hat H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a high ratio between chosen spins and flipped spins.
        Therefore it minimizes the calculations needed for a change of
        the state.

    \paragraph{Wolff}
        Close to the critical temperature \(T_c\) the Metropolis
        gets slower. This is called \emph{critical slowing} down and the
        cause is beyond the scope of this thesis.\\
        Using a cluster algorithm like the Wolff
        algorithm \cite{Wolff1989} speeds things up.
        For an Ising model the Wolff algorithm builds a cluster of sites
        starting with a random site and adding neighboring sites of the
        same spin with probability \(P_{\mathrm{add}}\) from eq. \eqref{eq:wolffAdd}
        \begin{equation}
            P_{\mathrm{add}} = 1-\exp\brac{-2\beta J}
            \label{eq:wolffAdd}
        \end{equation}
        Where \(J\) is the coupling constant (explained in section
        \ref{ssec:isingmodel}). The neighboring sites of the added sites
        are also considered and so forth. When there are no more sites
        to add, the spin of every site in the cluster is flipped
        \cite[S. ??]{NewmanBarkema1999} \cite[S. 151f]{Katzgraber2011}.
        This leads fast to new uncorrelated states at the critical
        temperature because big clusters are flipped. But there are not
        much advantages at high or low temperatures. At low temperatures
        the cluster will consist of almost all sites such that all but
        very few spins will be flipped. At high temperatures the cluster
        will only contain very few sites.
        Both situaitions have no advantage against the Metropolis algorithm.\\
        So one would activate this algorithm near the critical temperature
        but use a simple Metropolis algorithm at high and low temperatures.

    \paragraph{Parallel Tempering}
        %~ The main aim is to obtain the critical temperatures
        %~ \(T_c\) for different disorder paramters \(\sigma\).
        %~ Therefore it is necessary to simulate for many temperatures,
        %~ so that \emph{Parallel Tempering}\footnote{Before R. H.
            %~ Swendsen published this paper, a algorithm \(MC^3\) was
            %~ already published with the same idea. [citation needed]}
        %~ \cite{ParallelTempering1986} is a suited algorithm.
        Parallel Tempering simulates many identical systems at different
        temperatures and periodically swaps the spin configurations
        between two neighboring temperatures with probability \(P\) from
        eq. \eqref{eq:partemp} \cite[S. ??]{NewmanBarkema1999} \cite[S. 155ff]{Katzgraber2011}.
        \begin{equation}
            P((E_i,T_i) \to (E_{i+1},T_{i+1})) = \min\brac{1,\exp\brac{\brac{E_{i+1}-E_i}\brac{\frac{1}{T_{i+1}}-\frac{1}{T_i}}}}
            \label{eq:partemp}
        \end{equation}
        This has the advantages that correlation times of single
        temperatures are far smaller because their spin configuration
        gets often replaced by an other uncorrelated configuration. In
        many cases the more important advantage is, that a system which
        is trapped in a local minimum at a given temperature, can travel
        to higher temperatures, leave it's local minimum and cool down
        again in a lower minimum.\\
        In the case of a ferromagnetic Ising model the risk to get trapped
        in an local energy minimum is very low. Though the autocorrelation
        decreases significantly. In the scope of this thesis this comes
        for no cost, because one has to simulate for many temperatures
        to determine the critical temperature. The additional calculations
        to determine wheather to swap configurations or not are small in
        comparison with those that would be needed to generate a new
        uncorrelated state without \emph{parallel tempering}.

    \paragraph{Implementation Details}
        Here a mixture of the above three algorithms is used.
        Each sweep \(N\) Metropolis spin flips, one Wolff cluster flip
        and one parallel tempering swap are performed. Where \(N\) is the
        count of sites.\\
        Because it is not known before, where the critical temperatures
        \(T_c\) are located, the Wolff cluster algorithm is used for
        every temperature. The speed up at criticality is worth the
        moderate slow down at other temperatures.

\subsection{Equilibration- and Autocorrelation Time}
\label{ssec:eqtime}
    To generate an equilibrium state one starts with an arbitrary state
    and waits until it is equilibrated. The count of sweeps till
    equilibrium is called \emph{equilibration time} \(t_{eq}\).
    All measurement should start after this time.\\
    To determine when the system is in equilibrium, one can watch the
    development of some observables and take the point at which there
    are no big changes anymore as equilibrium. In fig.
    \ref{fig:equiandauto}\subref{sfig:equiandauto:equiE}
    the equilibrium is reached after approximately \(N=50\) sweeps for
    either an initial condition of all spins up and all spins random. It
    does not harm to double that value to be save.
    \begin{figure}[htbp]
        \centering
        \subfigure[Example of an equilibrating Ising system][]{
                \label{sfig:equiandauto:equiE}
                \includegraphics[width=0.47\textwidth]{plots/equiE}
        }
        \subfigure[Example of the autocorrelation of an Ising system][]{
                \label{sfig:equiandauto:autoM}
                \includegraphics[width=0.47\textwidth]{plots/autoM}
        }
        \caption[Examples for equilibration and autocorrelation]
        {
            \subref{sfig:equiandauto:equiE} Example of a Ising system
                \(L=64\) reaching equilibrium at \(T=2.36\) and
            \subref{sfig:equiandauto:autoM} the autocorrelation of an
                Ising system \(L=64\) at \(T=2.40\) (only Metropolis
                sweeps) on half logarithmic axis.
                The straight line is a exponential fit \(\exp(-t/\tau)\)
                with \(\tau = 342(1)\).
        }
        \label{fig:equiandauto}
    \end{figure}\\
    Because every state is generated from the state before, measurements
    of subsequent states are correlated. To determine when two states
    are independet, one calculates the normalized autocorrelation function
    \(\frac{\chi(t)}{\chi(0)}\), which should decay exponentially
    \(\chi(t) \propto \exp(t/\tau)\). This is visible in the half
    logarithmic plot \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}.
    To get the autocorrelation time one can either fit a exponential
    function \(\exp(-t/\tau)\) like in fig \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}
    or integrate \(\tau = \int \frac{\chi(t)}{\chi(0)} \de t\).
    \(\tau\) is an estimate after which time two samples are not
    correlated anymore. \cite[S. ??]{NewmanBarkema1999} \cite[S. 150f]{Katzgraber2011}.
    To make sure that the error is not underestimated one should wait
    \(2\tau\) sweeps between two measurements.
    The autocorrelation time is of course dependent on the temperature.
    At high temperatures the fluctuations are strong and subsequent
    states are more dissimilar and therefore less correlated than at low
    temperatures, where less spins are flipping. In this thesis always
    the longest autocorrelation time \(\tau\) of all simulated
    temperatures was taken.
    \[\tau = \max_T \tau_T\]

>>>>

\label{sec:montecarlo}
%~ Monte Carlo simulations are (probably) named after the casino in Monaco
%~ \cite{NewmanBarkema1999} because both rely on randomness.\\
The idea behind Monte Carlo simulations is to take random samples of
the observable, which should be measured, and estimate the observable from
this samples. In example the Monte Carlo integration chooses random
points within the bounding box of the fuction and estimates the integral
as the fraction of points below the function and points above muliplied
by the area of the bounding box.
Of course one wants to sample more points, where the area below the
function is big, because points in areas, where the function does
contribute little to the integral, contribute little to the integral.
This can be achieved by \emph{Importance Sampling}. One
generates random numbers according to some probabilty distribution, which
yields more numbers in the range where the function to ingegrate has high
values and weights the result accordingly.\\
%~ To improve this method: \emph{Importance Sampling}
%~ means that one takes more samples where the contribution of the
%~ function to the integral is bigger by using random numbers
%~ distributed........
The same principle is used to sample properties of statistical systems
by generating random states.
But it is difficult to create a random state of a physical system i.e. the
Ising system in equilibrium, hence one uses \emph{Markov Chains} to
generate new states \(\nu\) from old ones \(\mu\).
It is important that the transition probabilities \(A(\mu \to \nu)\)
obey \emph{Detailed Balance} and \emph{Ergodicity}.
\emph{Detailed Balance} means  that the probability to leave a state is
the same as the probability to enter the state in equilibrium
\(p_\mu A(\mu \to \nu) = p_\nu A(\nu \to \mu)\) with \(p_\mu\) the
probability to be in state \(\mu\).
And \emph{Ergodicity} requires that every possible state is reachable
from every other state in finite time. \cite{NewmanBarkema1999} \cite{Katzgraber2011}
Otherwise the samples might not be representative for the whole system.\\

\subsection{Monte Carlo Algorithms}
    \subsubsection{Metropolis}
        A Metropolis Monte Carlo\cite{Metropolis1953} simulation of an
        Ising model will choose a random spin, calculate the energy change
        \(\Delta \hat H\) defined in eq. \eqref{eq:dH} that would result
        from a flip of that spin and excute the flip with the probability \(A\)
        eq. \eqref{eq:metropolis} \cite{NewmanBarkema1999} \cite{Katzgraber2011}.
        \begin{equation}
            \Delta \hat H = \hat H(\nu) - \hat H(\mu)\\
            \label{eq:dH}
        \end{equation}
        \begin{equation}
            A(\mu \to \nu)
            \begin{cases}
                1                                 & \Delta \hat H \le 0 \\
                \exp{\brac{-\beta \Delta \hat H}} & \Delta \hat H > 0
            \end{cases}
            \label{eq:metropolis}
        \end{equation}
        So if a transition lowers the energy it is always done. This
        results in a high ratio between chosen spins and flipped spins.
        Therefore it minimizes the calculations needed for a change of
        the state.

    \subsubsection{Wolff}
        Close to the critical temperature \(T_c\) the Metropolis
        gets slower. This is called \emph{critical slowing} down and the
        cause is beyond the scope of this thesis.\\
        Using a cluster algorithm like the Wolff
        algorithm \cite{Wolff1989} speeds things up.
        For an Ising model the Wolff algorithm builds a cluster of sites
        starting with a random site and adding neighboring sites of the
        same spin with probability \(P_{\mathrm{add}}\) from eq. \eqref{eq:wolffAdd}
        \begin{equation}
            P_{\mathrm{add}} = 1-\exp\brac{-2\beta J}
            \label{eq:wolffAdd}
        \end{equation}
        Where \(J\) is the coupling constant (explained in section
        \ref{ssec:isingmodel}). The neighboring sites of the added sites
        are also considered and so forth. When there are no more sites
        to add, the spin of every site in the cluster is flipped
        \cite[S. ??]{NewmanBarkema1999} \cite[S. 151f]{Katzgraber2011}.
        This leads fast to new uncorrelated states at the critical
        temperature because big clusters are flipped. But there are not
        much advantages at high or low temperatures. At low temperatures
        the cluster will consist of almost all sites such that all but
        very few spins will be flipped. At high temperatures the cluster
        will only contain very few sites.
        Both situaitions have no advantage against the Metropolis algorithm.\\
        So one would activate this algorithm near the critical temperature
        but use a simple Metropolis algorithm at high and low temperatures.

    \subsubsection{Parallel Tempering}
        %~ The main aim is to obtain the critical temperatures
        %~ \(T_c\) for different disorder paramters \(\sigma\).
        %~ Therefore it is necessary to simulate for many temperatures,
        %~ so that \emph{Parallel Tempering}\footnote{Before R. H.
            %~ Swendsen published this paper, a algorithm \(MC^3\) was
            %~ already published with the same idea. [citation needed]}
        %~ \cite{ParallelTempering1986} is a suited algorithm.
        Parallel Tempering\cite{ParallelTempering1986} simulates many identical systems at different
        temperatures and periodically swaps the spin configurations
        between two neighboring temperatures with probability \(P\) from
        eq. \eqref{eq:partemp} \cite[S. ??]{NewmanBarkema1999} \cite[S. 155ff]{Katzgraber2011}.
        \begin{equation}
            P((E_i,T_i) \to (E_{i+1},T_{i+1})) = \min\brac{1,\exp\brac{\brac{E_{i+1}-E_i}\brac{\frac{1}{T_{i+1}}-\frac{1}{T_i}}}}
            \label{eq:partemp}
        \end{equation}
        This has the advantages that correlation times of single
        temperatures are far smaller because their spin configuration
        gets often replaced by an other uncorrelated configuration. In
        many cases the more important advantage is, that a system which
        is trapped in a local minimum at a given temperature, can travel
        to higher temperatures, leave it's local minimum and cool down
        again in a lower minimum.\\
        In the case of a ferromagnetic Ising model the risk to get trapped
        in an local energy minimum is very low. Though the autocorrelation
        decreases significantly. In the scope of this thesis this comes
        for no cost, because one has to simulate for many temperatures
        to determine the critical temperature. The additional calculations
        to determine wheather to swap configurations or not are small in
        comparison with those that would be needed to generate a new
        uncorrelated state without \emph{parallel tempering}.

    \subsubsection{Implementation Details}
        Here a mixture of the above three algorithms is used.
        Each sweep \(N\) Metropolis spin flips, one Wolff cluster flip
        and one parallel tempering swap are performed. Where \(N\) is the
        count of sites.\\
        Because it is not known before, where the critical temperatures
        \(T_c\) are located, the Wolff cluster algorithm is used for
        every temperature. The speed up at criticality is worth the
        moderate slow down at other temperatures.

\subsection{Equilibration- and Autocorrelation Time}
\label{ssec:eqtime}
    To generate an equilibrium state one starts with an arbitrary state
    and waits until it is equilibrated. The count of sweeps till
    equilibrium is called \emph{equilibration time} \(t_{eq}\).
    All measurement should start after this time.\\
    To determine when the system is in equilibrium, one can watch the
    development of some observables and take the point at which there
    are no big changes anymore as equilibrium. In fig.
    \ref{fig:equiandauto}\subref{sfig:equiandauto:equiE}
    the equilibrium is reached after approximately \(N=50\) sweeps for
    either an initial condition of all spins up and all spins random. It
    does not harm to double that value to be save.
    \begin{figure}[htbp]
        \centering
        \subfigure[Example of an equilibrating Ising system][]{
                \label{sfig:equiandauto:equiE}
                \includegraphics[width=0.47\textwidth]{plots/equiE}
        }
        \subfigure[Example of the autocorrelation of an Ising system][]{
                \label{sfig:equiandauto:autoM}
                \includegraphics[width=0.47\textwidth]{plots/autoM}
        }
        \caption[Examples for equilibration and autocorrelation]
        {
            \subref{sfig:equiandauto:equiE} Example of a Ising system
                \(L=64\) reaching equilibrium at \(T=2.36\) and
            \subref{sfig:equiandauto:autoM} the autocorrelation of an
                Ising system \(L=64\) at \(T=2.40\) (only Metropolis
                sweeps -- otherwise the decline is too steep to show)
                on half logarithmic axis.
                The straight line is a exponential fit \(\exp(-t/\tau)\)
                with \(\tau = 342(1)\).
        }
        \label{fig:equiandauto}
    \end{figure}\\
    Because every state is generated from the state before, measurements
    of subsequent states are correlated. To determine when two states
    are independet, one calculates the normalized autocorrelation function
    \(\frac{\chi(t)}{\chi(0)}\), which should decay exponentially
    \(\chi(t) \propto \exp(t/\tau)\). This is visible in the half
    logarithmic plot \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}.
    To get the autocorrelation time one can either fit a exponential
    function \(\exp(-t/\tau)\) like in fig \ref{fig:equiandauto}\subref{sfig:equiandauto:autoM}
    or integrate \(\tau = \int \frac{\chi(t)}{\chi(0)} \de t\).
    \(\tau\) is an estimate after which time two samples are not
    correlated anymore. \cite[S. ??]{NewmanBarkema1999} \cite[S. 150f]{Katzgraber2011}.
    To make sure that the error is not underestimated one should wait
    \(2\tau\) sweeps between two measurements.
    The autocorrelation time is of course dependent on the temperature.
    At high temperatures the fluctuations are strong and subsequent
    states are more dissimilar and therefore less correlated than at low
    temperatures, where less spins are flipping. But the longest
    autocorrelation times are encountered at the critical temperature.
    This effect is called \emph{critical slowing down} and is
    characterized by the \emph{dynamical critical exponent} \(z\)
    \cite{SwendsenWang1987}. The dependence of the autocorrelation time
    \(\tau\) on the system size \(L\) is then given by \(\tau \propto L^z\).

### 2013-06-28 Tex/chapters/results.tex
weitere Korrekturen und Umformulierungen

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}). The used values for different system sizes
are listed in tab. \ref{tab:tauAndTeq} in numbers of sweeps. Note that
the \(t_{eq}\) values are generously rounded up to be on the safe side
and the \(\tau\) values are determined as the maximum \(\tau\) over all
simulated temperatures.
\[\tau = \max_T \tau_T\]
\begin{table}[htbp]
    \center
    \begin{tabular}{l r r r r r}
        \toprule
        \(L\)    & 16 &  32 &  64 & 128 & 256\\
        \midrule
        \(\tau\) &  3 &   5 &   7 &  12 &  16\\
        \(t_eq\) & 40 & 100 & 100 & 200 & 600\\
        \bottomrule
    \end{tabular}
    \caption{Autocorrelation Times $\tau$ and Equilibration Times $t_{eq}$}
    \label{tab:tauAndTeq}
\end{table}\\
Obiviously the autocorrelation time \(\tau\) increases with the system
size. In fact it obeys a power law, which is the expected bahaivior as
mentioned in section \ref{ssec:eqtime}. In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw}
the dynamical critical exponent \(z\) is determined as \(z=0.57(4)\).
This exponent is much higher than the expectation. For the Wolff cluster
algorithm at criticality, one would expect \(z=0.25\)\cite{NewmanBarkema1999}.
This suggests that \emph{critical slowing down} is not an issue in this
case and the autocorrelation time is limited by the slow decorrelation
at low temperatures. A direct proof is given by fig. \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
Apparently even parallel tempering can not compensate that. Probably
because only one configuration swap between neighboring temperatures is
performed each sweep. Those neighboring configurations at low temperatures
have a magnetisation per site of \(m \approx \pm 1\), so there is a good
chance, that the exchange does not change \(m\) (in case of same sign)
and thus does not lower the autocorrelation function.
\begin{figure}[htbp]
    \centering
    \subfigure[in dependence on the system size $L$][]
    {
        \label{fig:autocorr:powerLaw}
        \includegraphics[width=0.45\textwidth]{plots/autocorrL}
    }
    \subfigure[in dependence on the temperature $T$][]
    {
        \label{fig:autocorr:temperatures}
        \includegraphics[width=0.45\textwidth]{plots/autocorrT}
    }
    \caption[The autocorrelation time $\tau$]
    {
        Dependece of the autocorrelation time $\tau$ on\\
        \subref{fig:autocorr:powerLaw} the system size $L$\\
        \subref{fig:autocorr:temperatures} the temperature $T$
    }
    \label{fig:autocorr}
\end{figure}\\
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(N_{\mathrm{measure}}=10000\) measurements for \(L=16,32\)
or \(N_{\mathrm{measure}}=5000\) for \(L=64,128\). The number of
calculated sweeps totals to
\[N_{\mathrm{sweeps}}=t_{eq}+2\tau N_{\mathrm{measure}}\]
The expected values \(\avg{O}\) for \(100\) different random proximity
graphs with the same disorder paramter \(\sigma\) are then averaged to
\(\overline{\avg{O}}\).
The errors are estimated by bootstrapping \cite{Bootstrap} with 200
bootstrap samples.

@@

    decline. The jump is only present in infinite systems, hence every
    computer simulation will show some \emph{finte size effects}.

>>>>

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}). The used values for different system sizes
are listed in tab. \ref{tab:tauAndTeq} in numbers of sweeps. Note that
the \(t_{eq}\) values are generously rounded up to be on the safe side
and the \(\tau\) values are determined as the maximum \(\tau\) over all
simulated temperatures \(\tau = \max_T \tau_T\) and rounded to the next
integer.
\begin{table}[htbp]
    \center
    \begin{tabular}{l r r r r r}
        \toprule
        \(L\)    & 16 &  32 &  64 & 128 & 256\\
        \midrule
        \(\tau\) &  3 &   5 &   7 &  12 &  16\\
        \(t_{eq}\) & 40 & 100 & 100 & 200 & 600\\
        \bottomrule
    \end{tabular}
    \caption{Autocorrelation Times $\tau$ and Equilibration Times $t_{eq}$}
    \label{tab:tauAndTeq}
\end{table}\\
Obiviously the autocorrelation time \(\tau\) increases with the system
size. In fact it obeys a power law, which is the expected bahaivior as
mentioned in section \ref{ssec:eqtime}. In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw}
the dynamical critical exponent \(z\) is determined as \(z=0.57(4)\).
This exponent is much higher than the expectation. For the Wolff cluster
algorithm at criticality, one would expect \(z=0.25\)\cite{NewmanBarkema1999}.
This suggests that \emph{critical slowing down} is not an issue in this
case and the autocorrelation time is limited by the slow decorrelation
at low temperatures. A direct proof is given by fig. \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
Apparently even parallel tempering can not compensate that. Probably
because only one configuration swap between neighboring temperatures is
performed each sweep. Those neighboring configurations at low temperatures
have a magnetisation per site of \(m \approx \pm 1\), so there is a good
chance, that the exchange does not change \(m\) (in case of same sign)
and thus does not lower the autocorrelation function.
\begin{figure}[htbp]
    \centering
    \subfigure[in dependence on the system size $L$][]
    {
        \label{fig:autocorr:powerLaw}
        \includegraphics[width=0.45\textwidth]{plots/autocorrL}
    }
    \subfigure[in dependence on the temperature $T$][]
    {
        \label{fig:autocorr:temperatures}
        \includegraphics[width=0.45\textwidth]{plots/autocorrT}
    }
    \caption[The autocorrelation time $\tau$]
    {
        Dependece of the autocorrelation time $\tau$ on\\
        \subref{fig:autocorr:powerLaw} the system size $L$\\
        \subref{fig:autocorr:temperatures} the temperature $T$
    }
    \label{fig:autocorr}
\end{figure}\\
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(N_{\mathrm{measure}}=10000\) measurements for \(L=16,32\)
or \(N_{\mathrm{measure}}=5000\) for \(L=64,128\). The number of
calculated sweeps totals to
\[N_{\mathrm{sweeps}}=t_{eq}+2\tau N_{\mathrm{measure}}\]
The expected values \(\avg{O}\) for \(100\) different random proximity
graphs with the same disorder paramter \(\sigma\) are then averaged to
\(\overline{\avg{O}}\).
The errors are estimated by bootstrapping \cite{Bootstrap} with 200
bootstrap samples.

@@

    decline. The jump is only present in infinite systems, hence no
    computer simulation will show the exact behaivior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
